---
title: "Evaluating the impact of modelling Omicron strain dynamics on short-term COVID-19 forecast performance"
subtitle: "Summary"
author: Sam Abbott, Katharine Sherratt, Sebastian Funk
bibliography: ../../writeup/library.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa-numeric-superscript-brackets.csl
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 4
    includes:
      before_body: ../header.html
      after_body: ../footer.html
---

```{r setup, echo = FALSE, cache = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, dpi = 330, warning = FALSE)
```

**This project is under active development. Current results are preliminary and may change as more data is collected or as the methodology is improved.**

# Background

```{r, child = "../chunks/introduction.Rmd"}
```

In this work, 

# Methods

We do x,y, and z. It is both very principled and very very impressive.

We give a brief int to our methods and explain where to find more.

We evaluate these models first visually across a range of forecasting dates and then quantitatively using proper scoring rules [@scoringutils] on both the natural and log scales (corresponding to absolute and relative performance) aggregating scores first across all targets and then stratifying in turn by location, horizon, date of postive test, and date of report. We also explore other aspects of our models performance by highlighting models that have problematic fitting diagnostics and summarising the estimation time for each model.

All models are implemented using the `forecast.vocs` R package [@forecast.vocs]. The forecasting and evaluation pipeline is implemented using the `targets` R package[@targets]. All input data, interim data, and output data are available and should also be fully reproducible from the provided code. Please see the resources section for details. Further details on our methodology are included in our [paper](https://epiforecasts.io/eval-omicron-for-case-forecasting/paper.pdf).

# Resources

```{r, child = "../chunks/documentation.Rmd"}
```

# References
