---
title: "Supplementary information: Evaluating the impact of modelling strain dynamics on short-term COVID-19 forecast performance"
output: 
  html_document:
    keep_md: false
    toc: true
    toc_float: true
    code_folding: hide
    theme: united
bibliography: library.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa-numeric-superscript-brackets.csl
header-includes:
   - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, eval = TRUE,
  fig.width = 9, fig.height = 9,
  dpi = 320,
  fig.path = "figures/"
)
```

## Authors

S. Abbott (1), K. Sherratt (1), S. Funk (1), ... (TBD)

Correspondence to: sam.abbott@lshtm.ac.uk

## Affiliations

1. Center for the Mathematical Modelling of Infectious Diseases, London School of Hygiene & Tropical Medicine, London WC1E 7HT, United Kingdom

## Overview

This analysis uses the `targets` package for reproducibility. The full analysis structure can be seen below:

```{r, message = FALSE}
library(targets)
tar_glimpse()
```

## Data

Notification data for Germany was sourced from ...
Sequence data for Germany was sourced from ...

Notification data for the United Kingdom was sourced from ...
Sequence data for the United Kingdom was sourced from ...

## Models

### Single strain

We model the mean ($\lambda_t$) of reported COVID-19 cases ($C_t$) as an order 1 autoregressive (AR(1)) process on the log scale by week ($t$). The model is initialised by assuming that the initial reported cases are representative with a small amount of error (2.5%).

\begin{align}
  \log \lambda_0 &\sim \text{LogNormal}\left(\log C_0 , 0.025 \times \log C_0 \right) \\
  \log \lambda_t &= \log \lambda_{t-1} + r_t
\end{align}

Where $r_t$ can be interpreted as the weekly growth rate and the exponential of $r_t$ as the effective reproduction number ($R_t$) assuming a generation time of 7 days. $r_t$ is then itself modelled as a differenced AR(1) process, 

\begin{align}
  r_0 &\sim \text{Normal}\left(0, 0.25 \right) \\
  r_t &= r_{t-1} + \epsilon_t  \\
  \epsilon_0 &= \eta_0 \\
  \epsilon_t &= \beta \epsilon_{t-1} + \eta_t
\end{align}

Where, 

\begin{align}
  \eta_t &\sim \text{Normal}\left(0, \sigma \right) \\
  \sigma &\sim \text{Normal}\left(0, 0.1 \right) \\
  \beta &\sim \text{Normal}\left(0, 1 \right)
\end{align}

We then assume a negative binomial observation model with overdispersion $\phi_c$ for reported cases ($C_t$),

\begin{align}
  C_{t} &\sim \text{NegBinomial}\left(\lambda_t, \phi_c\right) \\
  \frac{1}{\sqrt{\phi_c}} &\sim \text{Normal}(0, 1) 
\end{align}

Where $\sigma$, and $\frac{1}{\sqrt{phi_c}}$ are truncated to be greater than 0 and $\beta$ is truncated to be between -1 and 1.

The stan code for this model is as follows:

### Two strain

We model strain dynamics using the single strain model as a starting point but with the addition of strain specific AR(1) variation and a beta binomial observation process for sequence data. The full two strain model is described below. Parameters related to the delta variant are given the $\delta$ superscript and parameters related to non-delta cases are given the $o$ superscript.

Mean reported cases are again defined using a AR(1) process on the log scale for each strain and the combined for overall mean reported cases. 

\begin{align}
  \log \lambda_0 &\sim \text{LogNormal}\left(\log C_0 , 0.025 \times \log C_0 \right) \\
  \log \lambda^{\delta}_0 &\sim \text{LogNormal}\left(\log C^{\delta}_0 , 0.025 \times \log C^{\delta}_0 \right) \\
  \log \lambda^{o}_0 &= \log \left(\lambda_0 - \lambda^{\delta}_0 \right) \\
  \log \lambda^{\delta}_t &= \log \lambda^{\delta}_{t-1} + r^{\delta}_t \\
  \log \lambda^{o}_t &= \log \lambda^{o}_{t-1} + r^{o}_t \\
  \lambda_t &= \lambda^{\delta}_t + \lambda^{o}_t
\end{align}

Where $C^{\delta}_0$ is derived by calculating the mean proportion of cases that were delta for the first time point using the overall number of reported cases, the number of sequenced cases, and the number of sequences that were positive for the delta variant. The growth rate for delta and non-delta cases ($r^{o, \delta}_t$) is then modelled as a combination of an overall growth rate ($r_t$ as defined for the single strain model), a strain specific modifer ($s^{o, \delta}_0$), and an AR(1) error term post introduction ($\epsilon^{o, \delta}_t$).

\begin{align}
  r^{o, \delta}_t &= r_t + s^{o, \delta} + \epsilon^{o, \delta}_t \\
  \epsilon^{o, \delta}_0 &= 0 \\
  \epsilon^{o, \delta}_t &= \epsilon^{o, \delta}_{t-1} + \eta^{o, \delta}
\end{align}

Where, 

\begin{align}
  \eta^{o, \delta}_t &\sim \text{Normal}\left(0, \sigma^{o, \delta} \right) \\
  \sigma^{o, \delta} &\sim \text{Normal}\left(0, 0.1 \right) \\
  s^o &= 0 \\
  s^{\beta} &\sim \text{Normal}(0.2, 0.2)
\end{align}

Which assumes a relatively uninformed transmissibility advantage for the delta variant vs non-delta variant cases. The mean proportion of samples that have the delta variant ($p_t$) is then estimated using the mean reported cases with the delta variant and the overall mean reported cases.

\begin{equation} 
  p_t = \frac{\lambda^{\delta}_t}{\lambda_t}
\end{equation}

In addition to the negative binomial observation model for cases we assume a beta binomial observation model for the number of sequences ($N_t$) that are postive ($P_t$) for the delta variant each week with overdispersion $\phi_s$.

\begin{align}
  P_{t} &\sim \mathrm{BetaBinomial}\left(N_t, p_t \phi_s, (1 - p_t) \phi_s\right) \\
  \frac{1}{\sqrt{\phi_s}} &\sim \mathcal{N}(0, 1) 
\end{align}

Where $\sigma^{o, \delta}$, and $\frac{1}{\sqrt{\phi_s}}$ are truncated to be greater than 0.

The stan code for this model is follows

## Model evaluation

### Prior Predictive Checks

- Plot prior and observations for 4 forecast horizons in germany
- Plot model summmary statistics

#### Single strain

#### Two strain

### Posterior Predictive Checks

- Plot posterior and observations for 4 forecast horizons
- Plot model summmary statistics
- Plot fit statistics

#### Single strain

#### Two strain

##### Scaled

#### Pooled

#### Independent

### Simulation-based calibration

For 4 forecast dates draw 100 samples from the prior, fit the model, and report the coverage of the model fits for key parameters.

#### Single strain

#### Two strain

##### Scaled

##### Pooled

##### Independent

## Real-world evaluation

### Summary

- Plot from each case study showing two week ahead forecasts single + two strain models
- Table summarising scores for each case study
- Table summarising model fit statistics for each location

### Evaluation

## Data availability scenarios

### Scenarios

### Summary

### Evaluation

## Comparsion to prospective forecasts

## References

<div id = 'refs'></div>
